---
title: "Case Study 11"
author: Tina Ni
date: Novmember 12, 2024
output: github_document
---
```{r Load libraries}
library(tidyverse)
library(tidycensus)
library(spData)
library(sf)
library(mapview)
library(foreach)
library(doParallel)
library(leaflet)
registerDoParallel(4)
getDoParWorkers() # check registered cores
```

```{r install Census API key}
## census_api_key("YOUR CENSUS API KEY", install = TRUE, overwrite = TRUE)
```

```{r Get Block Level Data on 2020 Pop by Race in each Block in Buffalo.}
race_vars <- c(
  "Total Population" = "P1_001N",
  "White alone" = "P1_003N",
  "Black or African American alone" = "P1_004N",
  "American Indian and Alaska Native alone" = "P1_005N",
  "Asian alone" = "P1_006N",
  "Native Hawaiian and Other Pacific Islander alone" = "P1_007N",
  "Some Other Race alone" = "P1_008N",
  "Two or More Races" = "P1_009N"
)
options(tigris_use_cache = TRUE)

erie <- get_decennial (geography = "block", variables = race_vars, year = 2020, state = "NY", county = "Erie County", geometry = TRUE, sumfile = "pl", cache_table = TRUE )

```

```{r Crop county level data}
extent<- c(xmin=-78.9,xmax=-78.85,ymin=42.888,ymax=42.92)
erie_crop <- st_crop (erie, extent)
```
Write a foreach loop that does the following steps for each racial group in the variable column of the erie dataset and rbinds the results (e.g. .combine=rbind) into a single sf object.

You may want to convert the variable column into a factor and use levels() or use unique().

filter the the data to include only one race at time
use st_sample() to generate random points for each person that resided within each polygon.

If you use a pipe (%>%), you will have to set size=.$value.
The . indicates that the column comes from the dataset that was passed to the function.

See here for details on how to use the . in a pipe.

convert the points from st_sample() to spatial features with st_as_sf()

mutate to add a column named variable that is set to the current racial group (from the foreach loop)

```{r}
variables <- unique(erie_crop$variable)
variables

variables[1]
# variables <- variables[c(2:8)]
# variables

```

```{r}
matrix <- matrix(ncol = 0, nrow = 0)
df <- as.data.frame(matrix)
```
```{r}
# erie_crop2 <- matrix [0,0]

## test is the name because I was testing it out but at this point my computer screams at me and I don't want to run this code again to change the name of the dataset... 

test <- foreach(i = 1:8,.combine=rbind) %do% 
{
erie_crop %>%
filter(variable == variables[i]) %>%
st_sample(., size= .$value, type = "random") %>%
st_as_sf() %>%
mutate(v = variables[i])
}
unique(test$v)
```

```{r}
# t <- erie_crop %>%
#   filter(variable == variables[1]) %>%
#   filter(value != 0) %>%
#   st_sample(., size= .$value, "random") %>%
#   st_as_sf() %>%
#   mutate(v = variables[1])
```


```{r}
mapview(test, cex = 3)
```