---
title: "Case Study 05 - Beware the Canadians!"
author: "Tina Ni"
date: August 1, 2024
output: github_document
---
---
1. Generate a polygon that includes all land in NY that is within 10km of the Canadian border (not including the great lakes).
2. Calculate area in km^2. How much land will you need to defend from the Canadians?
---
---
- load packages
---
```{r Load Packages}
library(spData)
library(sf)
library(tidyverse)
library(units)
library(ggplot2)
```
---
- load and format data
---
``` {r edit the data}
data(world)
data("us_states")
albers <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"

Canada <- world |>
  filter (name_long == "Canada")

head(Canada)

NYS <- us_states |>
  filter(NAME == "New York")

str(NYS)
```
---
- transform NYS and Canada, and create buffer Canada
---
```{r Transform NYS and Canada, Buffer Canada}
tr_Canada <- st_transform(Canada,albers)

tr_NYS <- st_transform(NYS, albers)


buffer_C  <-st_buffer(tr_Canada, 10000)

```
---
- create a map of the buffer and nys area
---
``` {r create a map of the buffer and nys area}

i <- st_intersection (buffer_C, tr_NYS)

ggplot()+
  geom_sf(data = i)

plot <- (ggplot()+
geom_sf(data = tr_NYS)+
  geom_sf(data =i, fill = "red")+

labs (title = "New York Land within 10 km")+
  theme (plot.title = element_text(hjust = 0.5))
         )

plot
```
---
- convert st_area and make it into a string
---
```{r convert st_area and make it into a string}

area_m <- st_area (i)
area_km <- set_units(area_m, km^2)
str_num<- toString (area_km, "km")
km <- "km^2"
str_area <- paste(str_num,km)

print(str_area)

subtitle <- paste("Canadian buffer area within New York State:", str_area)
```
---
- add buffer numeric area on map plot
---
```{r Add buffer numeric area on map plot}
plot+
  labs(subtitle = subtitle)

```
